---
title: "day2 advent of code "
output: html_notebook
---
### part1 
our sub starts at both Position and height 0. 

we can only move forward but we can move down or up in depth. 
Hence we need to make sure if V1 == up we decrease the value of depth. 

Afterthat we can do a simple summarion of the forward movements and the depth movements (using group_buy) and multiply both numbers in the end. 

```{r libraries}
library(dplyr)
```


```{r data wrangling part 1}

# Forward position
forward.tibble<- as_tibble(forward)

forward <- filter(forward.tibble, V1 == 'forward') %>% summarise(sum(V2))


#Depth
up <- -1*filter(forward.tibble, V1 == 'up') %>% summarise(sum(V2))
down <- filter(forward.tibble, V1 == 'down') %>% summarise(sum(V2))

# calc
forward*sum(up, down)


# neat trick to look at it like that
table(forward.tibble)

```
### Part 2 

down X increases your aim by X units.
up X decreases your aim by X units.
forward X does two things:
It increases your horizontal position by X units.
It increases your depth by your aim multiplied by X.


```{r}
horizontal <- 0
depth <- 0
aim <- 0
for(i in 1:length(forward.tibble$V1)){
  direction <- forward.tibble$V1[i]
  distance <- forward.tibble$V2[i]
  if(direction=="forward"){
    horizontal <- horizontal+distance
    depth <- depth+(aim*distance)
  } else if(direction=="up"){
    aim <- aim-distance
  } else if(direction=="down"){
    aim <- aim+distance
  }
}
print(horizontal*depth)
```
```{r part 2 adding the aim}
mutate(forward.tibble, aim = ifelse(V1 == 'up', -1*V2, V2))

```
