---
title: "Day3"
output: html_notebook
---
#Part 1

first i split the binary numbers in different columns using Excel. 
This can be done in R using tidyr::separate(data, c("1"..."12")) but its so easy in excel i preffer it :D
```{r libraries}
library(tidyverse)
```

```{r first attempt very very dirty and bad}
input<-select(input, -Column1.13)


table(input$Column1.1)
table(input$Column1.2)
#do this for 12 columns and manually check what's the most common digit in each 

```

for the binary number I used this [website](https://www.rapidtables.com/convert/number/binary-to-decimal.html)

#Part 2

very neat answer i found on [reddit](https://www.reddit.com/r/adventofcode/comments/r7r0ff/comment/hn1kd75/?utm_source=share&utm_medium=web2x&context=3) 

Also note neater way of what i did in excel `data03 <- do.call(rbind, strsplit(readLines("Day3.txt"), ""))`

```{r part2}
data03 <- do.call(rbind, strsplit(readLines("Day3.txt"), ""))
    find_c_o2 <- function(o2 = TRUE) {
      res <- data03
      for (k in seq_along(data03[1, ])) {
         x <- as.integer(res[,k])
         idx <- which(x == (if (o2) 1L else 0L))
         res <- if (sum(x) >= length(x) / 2)  res[idx, ] else res[-idx,]
         if (length(res) == length(data03[1, ])) break
      }
      return(strtoi(paste(res, collapse = ""), base = 2))
    }
    
    find_c_o2(TRUE) * find_c_o2(FALSE)
```


